---
title: "MTT_study3_cleaning2"
author: "Jen Richmond"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Phenomenological Data

This Rmd file reads in, cleans up and write to csv phenomenological data for the MTT project. 

# load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse) # data wrangling/visualisation
library(here) # for avoiding filepath drama
library(readxl) # for reading excel sheets
library(janitor) # for cleaning functions
library(naniar) # visualising missing data
library(wesanderson) #colour palettes
library(ggbeeswarm)


```

# read data

The R compatible version of the raw phenomological data lives in the xls sheet called all_clinical_data_Phenon_R within the sheet called Study3_MTT_Clinical_group. Use readxl to read in just that sheet

```{r}
phenom <- readxl::read_excel(here("data", "Study 3_MTT_Clinical Group_ Data (65 or under).xlsx"), sheet = "all_clinical_data_Phenom_R") %>%
  clean_names() 

```

# renaming

This is stupidly long code that renames all 185 variables to make the naming conventions consistent so that it will work with pivot_longer. 

```{r}
phenom_renamed <- phenom %>%
  rename(past_neutral_e1_text = past_neutral_6, 
         past_neutral_e1_vivid = vivid_7, 
         past_neutral_e1_prere = pre_re_8, 
         past_neutral_e1_sense1 = sense_9, 
         past_neutral_e1_sense2 = sense_10, 
         past_neutral_e1_context1 = context_11, 
         past_neutral_e1_context2 = context_12, 
         past_neutral_e1_perspective1 = perspective_13, 
         past_neutral_e1_perspective2 = perspective_14, 
         past_neutral_e1_ratio = ratio_15, 
         past_neutral_e1_intensity = intensity_16, 
         past_neutral_e1_reaction = reaction_17, 
         past_neutral_e1_impt = importance_18,
         past_neutral_e1_coh = coherence_19, 
         past_neutral_e1_temp = temporal_20, 
         past_neutral_e2_text = past_neutral_21, 
         past_neutral_e2_vivid = vivid_22, 
         past_neutral_e2_prere = pre_re_23, 
         past_neutral_e2_sense1 = sense_24, 
         past_neutral_e2_sense2 = sense_25, 
         past_neutral_e2_context1 = context_26, 
         past_neutral_e2_context2 = context_27, 
         past_neutral_e2_perspective1 = perspective_28, 
         past_neutral_e2_perspective2 = perspective_29, 
         past_neutral_e2_ratio = ratio_30, 
         past_neutral_e2_intensity = intensity_31, 
         past_neutral_e2_reaction = reaction_32, 
         past_neutral_e2_impt = importance_33,
         past_neutral_e2_coh = coherence_34, 
         past_neutral_e2_temp = temporal_35,
        past_negative_e1_text = past_negative_36, 
         past_negative_e1_vivid = vivid_37, 
         past_negative_e1_prere = pre_re_38, 
         past_negative_e1_sense1 = sense_39, 
         past_negative_e1_sense2 = sense_40, 
         past_negative_e1_context1 = context_41, 
         past_negative_e1_context2 = context_42, 
         past_negative_e1_perspective1 = perspective_43, 
         past_negative_e1_perspective2 = perspective_44, 
         past_negative_e1_ratio = ratio_45, 
         past_negative_e1_intensity = intensity_46, 
         past_negative_e1_reaction = reaction_47, 
         past_negative_e1_impt = importance_48,
         past_negative_e1_coh = coherence_49, 
         past_negative_e1_temp = temporal_50, 
         past_negative_e2_text = past_negative_51, 
         past_negative_e2_vivid = vivid_52, 
         past_negative_e2_prere = pre_re_53, 
         past_negative_e2_sense1 = sense_54, 
         past_negative_e2_sense2 = sense_55, 
         past_negative_e2_context1 = context_56, 
         past_negative_e2_context2 = context_57, 
         past_negative_e2_perspective1 = perspective_58, 
         past_negative_e2_perspective2 = perspective_59, 
         past_negative_e2_ratio = ratio_60, 
         past_negative_e2_intensity = intensity_61, 
         past_negative_e2_reaction = reaction_62, 
         past_negative_e2_impt = importance_63,
         past_negative_e2_coh = coherence_64, 
         past_negative_e2_temp = temporal_65, 
        past_positive_e1_text = past_positive_66, 
         past_positive_e1_vivid = vivid_67, 
         past_positive_e1_prere = pre_re_68, 
         past_positive_e1_sense1 = sense_69, 
         past_positive_e1_sense2 = sense_70, 
         past_positive_e1_context1 = context_71, 
         past_positive_e1_context2 = context_72, 
         past_positive_e1_perspective1 = perspective_73, 
         past_positive_e1_perspective2 = perspective_74, 
         past_positive_e1_ratio = ratio_75, 
         past_positive_e1_intensity = intensity_76, 
         past_positive_e1_reaction = reaction_77, 
         past_positive_e1_impt = importance_78,
         past_positive_e1_coh = coherence_79, 
         past_positive_e1_temp = temporal_80, 
         past_positive_e2_text = past_positive_81, 
         past_positive_e2_vivid = vivid_82, 
         past_positive_e2_prere = pre_re_83, 
         past_positive_e2_sense1 = sense_84, 
         past_positive_e2_sense2 = sense_85, 
         past_positive_e2_context1 = context_86, 
         past_positive_e2_context2 = context_87, 
         past_positive_e2_perspective1 = perspective_88, 
         past_positive_e2_perspective2 = perspective_89, 
         past_positive_e2_ratio = ratio_90, 
         past_positive_e2_intensity = intensity_91, 
         past_positive_e2_reaction = reaction_92, 
         past_positive_e2_impt = importance_93,
         past_positive_e2_coh = coherence_94, 
         past_positive_e2_temp = temporal_95, 
        future_neutral_e1_text = future_neutral_96, 
         future_neutral_e1_vivid = vivid_97, 
         future_neutral_e1_prere = pre_re_98, 
         future_neutral_e1_sense1 = sense_99, 
         future_neutral_e1_sense2 = sense_100, 
         future_neutral_e1_context1 = context_101, 
         future_neutral_e1_context2 = context_102, 
         future_neutral_e1_perspective1 = perspective_103, 
         future_neutral_e1_perspective2 = perspective_104, 
         future_neutral_e1_ratio = ratio_105, 
         future_neutral_e1_intensity = intensity_106, 
         future_neutral_e1_reaction = reaction_107, 
         future_neutral_e1_impt = importance_108,
         future_neutral_e1_coh = coherence_109, 
         future_neutral_e1_temp = temporal_110, 
         future_neutral_e2_text = future_neutral_111, 
         future_neutral_e2_vivid = vivid_112, 
         future_neutral_e2_prere = pre_re_113, 
         future_neutral_e2_sense1 = sense_114, 
         future_neutral_e2_sense2 = sense_115, 
         future_neutral_e2_context1 = context_116, 
         future_neutral_e2_context2 = context_117, 
         future_neutral_e2_perspective1 = perspective_118, 
         future_neutral_e2_perspective2 = perspective_119, 
         future_neutral_e2_ratio = ratio_120, 
         future_neutral_e2_intensity = intensity_121, 
         future_neutral_e2_reaction = reaction_122, 
         future_neutral_e2_impt = importance_123,
         future_neutral_e2_coh = coherence_124, 
         future_neutral_e2_temp = temporal_125, 
        future_negative_e1_text = future_negative_126, 
         future_negative_e1_vivid = vivid_127, 
         future_negative_e1_prere = pre_re_128, 
         future_negative_e1_sense1 = sense_129, 
         future_negative_e1_sense2 = sense_130, 
         future_negative_e1_context1 = context_131, 
         future_negative_e1_context2 = context_132, 
         future_negative_e1_perspective1 = perspective_133, 
         future_negative_e1_perspective2 = perspective_134, 
         future_negative_e1_ratio = ratio_135, 
         future_negative_e1_intensity = intensity_136, 
         future_negative_e1_reaction = reaction_137, 
         future_negative_e1_impt = importance_138,
         future_negative_e1_coh = coherence_139, 
         future_negative_e1_temp = temporal_140, 
         future_negative_e2_text = future_negative_141, 
         future_negative_e2_vivid = vivid_142, 
         future_negative_e2_prere = pre_re_143, 
         future_negative_e2_sense1 = sense_144, 
         future_negative_e2_sense2 = sense_145, 
         future_negative_e2_context1 = context_146, 
         future_negative_e2_context2 = context_147, 
         future_negative_e2_perspective1 = perspective_148, 
         future_negative_e2_perspective2 = perspective_149, 
         future_negative_e2_ratio = ratio_150, 
         future_negative_e2_intensity = intensity_151, 
         future_negative_e2_reaction = reaction_152, 
         future_negative_e2_impt = importance_153,
         future_negative_e2_coh = coherence_154, 
         future_negative_e2_temp = temporal_155, 
        future_positive_e1_text = future_positive, 
         future_positive_e1_vivid = vivid_157, 
         future_positive_e1_prere = pre_re_158, 
         future_positive_e1_sense1 = sense_159, 
         future_positive_e1_sense2 = sense_160, 
         future_positive_e1_context1 = context_161, 
         future_positive_e1_context2 = context_162, 
         future_positive_e1_perspective1 = perspective_163, 
         future_positive_e1_perspective2 = perspective_164, 
         future_positive_e1_ratio = ratio_165, 
         future_positive_e1_intensity = intensity_166, 
         future_positive_e1_reaction = reaction_167, 
         future_positive_e1_impt = importance_168,
         future_positive_e1_coh = coherence_169, 
         future_positive_e1_temp = temporal_170, 
         future_positive_e2_text = future_positive_2, 
         future_positive_e2_vivid = vivid_172, 
         future_positive_e2_prere = pre_re_173, 
         future_positive_e2_sense1 = sense_174, 
         future_positive_e2_sense2 = sense_175, 
         future_positive_e2_context1 = context_176, 
         future_positive_e2_context2 = context_177, 
         future_positive_e2_perspective1 = perspective_178, 
         future_positive_e2_perspective2 = perspective_179, 
         future_positive_e2_ratio = ratio_180, 
         future_positive_e2_intensity = intensity_181, 
         future_positive_e2_reaction = reaction_182, 
         future_positive_e2_impt = importance_183,
         future_positive_e2_coh = coherence_184, 
         future_positive_e2_temp = temporal_185)
```

separate text from ratings

```{r}
event_text <- phenom_renamed %>%
  select(1:5, ends_with("text")) 

  event_text %>% write_csv(here("data", "study3_event_text.csv"))

phenom_ratings <- phenom_renamed %>%
  select(1:5, !ends_with("text"))

```

# make ratings long
```{r}
phenom_long <- phenom_ratings %>%
  pivot_longer(names_to = c("direction", "valence", "event", "rating"), 
               values_to = "score", 
               names_sep = "_", 
               6:173)

phenom_long %>% write_csv(here("data", "study3_phenom.csv"))
```

# Details re phenom analysis

- vividness
- pre/re-experiencing
- sensory details
  + individual sensory items (sight, sound, smell, taste) averaged into single sensoty detail rating for each event
- contextual details
  + ratings of clarity of location and clarity of spatial arrangement of objects averaged into combined contextual details rating for each event
- perspective
  + ratio calculated by dividing rating for "seen through my own eyes" and "as if I was a fly on the wall" to determine field or observer perspective. Higher ratios indicate field perspecitve; lower ratios field perspective
- reaction
  + combined reaction score calculated by averaging intensity and physical reactions
- coherence
- importance
  
All phenomenological ratings averaged across event type (i.e. future neutral, future positive)





