---
title: "plotting_specficity"
author: "Jen Richmond"
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

```{r loadpackages, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(ggeasy)
library(here)
library(ggbeeswarm)
```

# read data

```{r readdata}
mtt_long <- read_csv(here("data", "study3_clean_mtt.csv"))
```

# SPECIFICITY

Each participant was asked to talk about 2 events in each condition category

- past, neutral
- past, negative
- past, positive
- future, neutral
- future, negative
- future, positive

These 12 events were coded as 1 = general, 2 = categorical, 3 = specific. Only specific events can be coded for internal vs. external details. 

Filter out events that weren't coded as specific.

```{r getspecifictyscores, message=FALSE, warning=FALSE}
specificity <- mtt_long %>%
  filter(code == 3)

#fix order of group factor
specificity$group <- fct_relevel(specificity$group, c("CON", "ANX"))
                                                     
```

# Replicating plots from draft paper

The original analysis looked at proportion internal by direction (past/future), valence (pos, neutral, neg), group (anx, con), finding only a main effect of group.  

### 1. Main effect of group 

```{r overgendirection, message=FALSE, warning=FALSE}
specificity %>%
  group_by(group) %>%
  summarise(mean = mean(proportion, na.rm = TRUE), 
            sd = sd(proportion, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = group, y = mean, fill = group)) +
  geom_col() +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
  scale_fill_brewer(palette="Dark2") +
  ylim(0,1) +
  labs(title = "Proportion internal by group", subtitle = "averaged across valence and direction")
```

