---
title: "MTT_study3_cleaning"
author: "Jen Richmond"
date: "27/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

```{r message=FALSE, warning=FALSE}
library(tidyverse) # data wrangling/visualisation
library(here) # for avoiding filepath drama
library(readxl) # for reading excel sheets
library(janitor) # for cleaning functions
library(naniar) # visualising missing data
library(wesanderson) #colour palettes
library(ggbeeswarm)
```

# read data

The R compatible version of the raw data lives in the xls sheet called Detail_Analysis_R within the sheet called Study3_MTT_Clinical_group. Use readxl to read in just that sheet

```{r}
mtt <- readxl::read_excel(here("data", "Study 3_MTT_Clinical Group_ Data (65 or under).xlsx"), sheet = "Detail_Analysis_R") %>%
  clean_names() 

table(mtt$group, mtt$identifier)

# filter only pp who 
mtt_complete <- mtt %>%
  filter(identifier == "Good")


```


# clean data

### make it long

```{r}
# list variable names
names(mtt_complete)

# make long to get columns for direction, valence, event etc 
mtt_longest <- mtt_complete %>%
  pivot_longer(names_to = c("direction", "valence", "event", "measure"), 
               values_to = "response", 
               names_sep = "_", future_neutral_e1_code:future_neg_e2_proportion) %>%
  mutate(id = row_number())

# make wide to have dvs in different columns, will throw a warning about duplicates /list
mtt_long <- mtt_longest %>%
  pivot_wider(names_from = measure, values_from = response, id_cols = px_no:event) %>%
  arrange(px_no, direction, valence, event)


```

notes: 

- pp_no problems, there are two pp1021, renamed in excel as 1021J and 1021S

### fix list issue (NO LONGER AN ISSUE)

Trying to work out why variables code through proportion are lists.  

Ahhh this whole list issue was caused by duplicated participant numbers. The chunk below is probably not needed anymore
```{r}
mtt_long_df <- as.data.frame(mtt_long)

glimpse(mtt_long_df) # so weird that code:proportion are list type, ridiculous work around required because I don't really GET lists

```


```{r eval = FALSE}
# pull code, internal, external, and proportion into values in the environment using unlist

codes <- unlist(mtt_long$code) # weird that this has 516 obs not 504
internal <- unlist(mtt_long$internal)
external <- unlist(mtt_long$external)
proportion <- unlist(mtt_long$proportion)

# use col bind to join them and then make into a data frame
values <- cbind(codes, internal, external, proportion)
values <- as.data.frame(values) 

#drop the extra empty rows at the bottom to get 504 obs again

mtt_values <- values[-c(505:516), ] 

# select just the variables from the mtt_df
mtt_variables <- mtt_long_df %>%
  select(1:6)

#join the mtt variables and values using cbind

mtt_done <- cbind(mtt_variables, mtt_values)

#did that fix the data type issue?

glimpse(mtt_done)

# YES
```

### check proportion calc

The proportions were calculated in excel, lets just make sure they are correct by making a new variable called prop, and comparing it to the imported proportion variable

```{r}
mtt_check <- mtt_long %>%
  mutate(prop = internal/(internal + external)) %>%
  mutate(check = if_else(prop == proportion, TRUE, FALSE))

# check missingness

vis_miss(mtt_long)

```
That looks ok, only a few NAs in the code column, more in the int/ext/prop columns because those will be NA if code is not = 3. 

### write cleaned to csv

```{r}
mtt_long %>%
  write_csv(here("data", "study3_clean_mtt.csv"))
```









