---
title: "phenomological rating plots"
author: "Jen Richmond"
date: "09/09/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(ggeasy)
```

# read in data
```{r message=FALSE, warning=FALSE}
phenom <- read_csv(here("data", "study3_phenom.csv"))
```

Filter only pp who are good, fix data types
```{r}

phenom_complete <- phenom %>%
  filter(identifier == "Good")  %>%
  arrange(px_no, valence, direction, event)


phenom_complete <- phenom_complete %>%
  mutate_if(is.character, as.factor)

phenom_complete$valence <- fct_relevel(phenom_complete$valence, 
                    c("positive", "neutral", "negative"))

phenom_complete$group <- fct_relevel(phenom_complete$group, 
                                       c("CON", "ANX"))

phenom_complete$direction <- fct_relevel(phenom_complete$direction, 
                                       c("past", "future"))

```


# Phenomolgical variables

- vividness
- pre/re-experiencing
- sensory details
  + individual sensory items (sight, sound, smell, taste) averaged into single sensoty detail rating for each event
- contextual details
  + ratings of clarity of location and clarity of spatial arrangement of objects averaged into combined contextual details rating for each event
- perspective
  + ratio calculated by dividing rating for "seen through my own eyes" and "as if I was a fly on the wall" to determine field or observer perspective. Higher ratios indicate field perspecitve; lower ratios field perspective
- reaction
  + combined reaction score calculated by averaging intensity and physical reactions
- coherence
- importance
  
All phenomenological ratings averaged across event type (i.e. future neutral, future positive)

# Plots

# VIVIDNESS 
```{r}
phenom_complete %>%
  filter(rating == "vivid") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
 geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated vividness")
```

# PRE-RE experiencing
```{r}
phenom_complete %>%
  filter(rating == "prere") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
 geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated pre/re experiencing")
```

# SENSORY
```{r}
phenom_complete %>%
  filter(rating %in% c("sense1", "sense2")) %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
 geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated sensory")
```

# CONTEXT
```{r}
phenom_complete %>%
  filter(rating %in% c("context1", "context2")) %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
 geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated context")
```

# PERSPECTIVE
```{r}
phenom_complete %>%
  filter(rating == "ratio") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
 geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated perspective ratio", subtitle = "ratios above 1 mean that event was seen more through own eyes than as fly on wall", caption = "higher = field perspective, lower = observer perspective")
```

Check calculations on perspective ratio

```{r}
perspective <- phenom_complete %>%
  filter(rating %in% c("perspective1", "perspective2", "ratio"))

persp_wide <- perspective %>%
  pivot_wider(names_from = rating, 
              values_from = score) %>%
  mutate(ratio_test = perspective1/perspective2)
```

Yup calculation seems fine. 

# REACTION

```{r}
phenom_complete %>%
  filter(rating == "reaction") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated reaction (intensity + physical reaction")
```

# COHERENCE

### group x direction x valence
```{r}
phenom_complete %>%
  filter(rating == "coh") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated coherence")
```

### just group effect
```{r}
phenom_complete %>%
  filter(rating == "coh") %>%
  group_by(group) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = group, y = mean, fill = group)) +
  geom_col() +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
  ylim(0, 7) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated coherence", subtitle = "people with anxiety rate events as more (not less) coherent")
```

# IMPORTANCE

### group x direction x valence
```{r}
phenom_complete %>%
  filter(rating == "impt") %>%
  group_by(group, direction, valence) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = valence, y = mean, fill = valence)) +
  geom_col() +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
 facet_grid(vars(group), vars(direction)) +
  ylim(0, 6) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated importance")

```

group x direction interaction 
```{r}

phenom_complete %>%
  filter(rating == "impt") %>%
  group_by(group, direction) %>%
  summarise(mean = mean(score, na.rm = TRUE), 
            sd = sd(score, na.rm = TRUE), 
            n = n(), 
            stderr = sd/sqrt(n)) %>%
  ggplot(aes(x = group, y = mean, fill = direction)) +
  geom_col(position = "dodge") +
  geom_errorbar(aes(ymin=mean-stderr, ymax=mean+stderr),
                  size=.3,    # Thinner lines
                    width=.2,
                      position=position_dodge(.9)) +
  ylim(0,6) +
  scale_fill_brewer(palette="Dark2") +
  labs(title = "Rated importance", subtitle = "people with anxiety rate future events as more important")


```